```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

> "There are no routine statistical question; only questionable statistical routines." 
>
> --- **D.R. Cox** quoted in Chatfield, C. 1991. Avoiding statistical pitfalls. Statistical Science 6: 240-268.


# Exercises

## (1 Point) Exercise 0 (Students To the GitHub Mobile!)

Keeping in line with our commitment to using versioning control, please:

1. Initialize a new repository on GitHub called `stat385hw2` 
2. Create a new RStudio project for Homework 2. 
3. **Please commit every exercise as you complete them.**
4. Provide a link to your GitHub repository to be awarded credit.

![figure1.png]

## (4 Points) Exercise 1 (It's a Mad, Mad, Mad, Mad Computational World)



The tenets behind this exercise are to prepare you to create different implementations and provide benchmarks for each implementation. 

Recall the different Sum of Squares, (*RSS*, *FSS*, *TSS*) formulations presented in [Lecture 5](http://stat385.thecoatlessprofessor.com/assets/lectures/lec5/lec5_functions_recursion_memoization_benchmarking.pdf#page=18). 

- Recreate the three sum of square functions given using loops **AND** recreate `tss_relationship` such that it uses these newly created functions.
   - Thus, you must "unvectorize" the computation. 
   - Call these functions: `compute_rss_loop`, `compute_fss_loop`, `compute_tss_loop`, and `tss_relationship_loop`.


```{r}
n = 1e4
x = seq(0, 1, length.out = n)
set.seed(123)
y = runif(n)
y_bar = mean(y)
y_hat = lm(y~x)$fitted.values

#rss loop
compute_rss_loop = function(y, y_hat) {
rss_element = rep(0, n)
for (i in 1:n) {
  rss_element[i] = (y[i] - y_hat[i])^2
  }
return(sum(rss_element))
}

#fss loop
compute_fss_loop = function(y_hat, y_bar) {
fss_element = rep(0, n)
for (i in 1:n) {
  fss_element[i] = (y_hat[i] - y_bar)^2
  }
return(sum(fss_element))
}

#tss loop
compute_tss_loop = function(y, y_bar) {
tss_element = rep(0, n)
for (i in 1:n) {
  tss_element[i] = (y[i] - y_bar)^2
  }
return(sum(tss_element))
}

tss_relationship_loop = function(y, y_hat, y_bar) {
  rss = compute_rss_loop(y, y_hat)
  fss = compute_fss_loop(y_hat, y_bar)
  tss = rss + fss
  return(tss)
}

```


- Check that the looping functions are equivalent to the vectorized components via `all.equal()`. Does the equality change between functions if the `tolerance` parameter in `all.equal` changes? (e.g. from `1e-3` to `0`) **Why?**
    - Use the [Setup Function Call](http://stat385.thecoatlessprofessor.com/assets/lectures/lec5/lec5_functions_recursion_memoization_benchmarking.pdf#page=22) to generate data.
    - Increase the number of observation in the script to `n = 1e4`
    - **Note:** You will have to use `check.attributes = F` to avoid `"names for current but not for target"`

```{r}
rss = compute_rss_loop(y, y_hat)
fss = compute_fss_loop(y_hat, y_bar)
tss = compute_tss_loop(y, y_bar)

rss_vector = function(y, y_hat) {
return(sum((y - y_hat)^2))
}


fss_vector = function(y_hat, y_bar) {
  return(sum((y_hat - y_bar)^2))
}

tss_vector = function(y, y_bar) {
  return(sum((y - y_bar)^2))
}

rss_v2 = rss_vector(y, y_hat)
fss_v2 = fss_vector(y_hat, y_bar)
tss_v2 = tss_vector(y, y_bar)

all.equal(rss,rss_v2)
all.equal(fss,fss_v2)
all.equal(tss,tss_v2)

```

```{r}
all.equal(rss,rss_v2, tolerance = 1e-3)
all.equal(fss,fss_v2, tolerance = 1e-3)
all.equal(tss,tss_v2, tolerance = 1e-3)

all.equal(rss,rss_v2, tolerance = 0)
all.equal(fss,fss_v2, tolerance = 0)
all.equal(tss,tss_v2, tolerance = 0)
```

**Even I changed the tolerance of `all.equal` the results stayed the same, because the two functions (from loops or vector) are identical.**

- Perform the two different types of benchmarking operations using `rbenchmark` and `microbenchmark` as discussed in class. Comment as to whether there is a noticably difference in performance between the functions. Provide a recommendation as to the correct style to use in the future.
    - Make sure to **cache** the benchmark results in the `Rmd` document!

```{r, cache = T}
install.packages("rbenchmark")
```

```{r, cache = T}
library('rbenchmark')
out1 = benchmark(compute_rss_loop(y, y_hat), rss_vector(y, y_hat),
columns = c("test", "replications", "elapsed", "relative"),
order="relative", 
replications=20) 
out1
```

```{r}
install.packages("microbenchmark")
```

```{r, cache= T}
library('microbenchmark')
out2 = microbenchmark(compute_rss_loop(y, y_hat), rss_vector(y, y_hat), times = 20)
summary(out2)
```

**The `loop` function spends much more time to run, so the `vector` function should be used.**

## (5 Points) Exercise 2 (Awake in a Nightmare)

The objective in this exercise is to understand unfamiliar code and improve it. While working on this problem, keep in mind the following quotation:

> "Always code as if the guy who ends up maintaining your code will be a violent psychopath who knows where you live. Code for readability."
> 
> --- John F. Woods ([Newsgroup comp.lang.c++](https://groups.google.com/forum/#!msg/comp.lang.c++/rYCO5yn4lXw/oITtSkZOtoUJ))


* Recently, a coworker left the company and your boss asked you to figure out the computations behind a graph the coworker had written. He wasn't entirely sure what was happening behind the scenes but the graph looked nice and the code could still be run. Your previous co-worker wrote the following code:

```{r, cache = T}
d = function(x){  length(x) }

s = function(x){ x*x }

u = function(x, ...){m = 0; k = 1; while(k <= d(x)){ m = m + x[k]; k = k + 1; }; m/d(x)}

w = function(b, v, q = T){n = d(b); o = 0; for(i in seq_len(n)){ o = o + s(b[i]-v) }; (1/(n-1*(q == T))*o)^(0.5)}

n = function(p, i, ...){ m = u(i, ...); r = w(i, m); (p-m)/r}

z = function(p, i, h = T, ...){
  g = seq(-3,3,length=1000);gg = dnorm(g)
  plot(g, gg, type="l", lwd=1, ...)
  k = n(p = p,i = i, ...)
  if(h) { tt = c(g[g<=k],k,-3); ll = c(gg[g<=k],0,0) }else{ tt = c(k,g[g>=k],3); ll = c(0,gg[g>=k],0) }
  polygon(tt, ll, col="red")
}
```

* The coworker left the following comment:

```{r coworker, cache = T, eval = F}
# Main function runs well
i = c(-1.21,0.28,1.08,-2.35,0.43,0.51,-0.57,-0.55,-0.56,-0.89)
z(1.57,i)
z(-2.33,i,h=F)
```

1. First try to figure out what is happening 
    - What distribution does the graph look like?
    *normal distribution*
    - What kind of calculation is occurring?
    **`d` gets the length of `x`.**
    **`s` gets $x^2$.**
    **`u` gets the mean of `x`.**
    **`w` gets the unbiased estimator of SD of `x`.**
    **`n` gets the standardized value of `x`.**
    **`z` gets the graph**
    - Are there any redundancies in the code? (e.g. redefinition of common functions)
2. Document each part of the code using comments `# Function does...`
    - Provide clarity behind the parameters of the functions.
    - Emphasize what is being returned at each stage. 
    - You may need to reference help for function information via: `?`
3. Try to clean up the code. 
    - Use intuitive variable and function names
    - Correct the indentation and split statements up to multiple lines.
    - Implement vectorized or base function.
**the cleaned codes**
```{r, cache = T}
d = function(x){  
  length(x) 
}                       # length of x

s = function(x){ 
  x*x 
}                       # the square of x

u = function(x, ...){
  m = 0
  k = 1
  while(k <= d(x)){ 
    m = m + x[k]
    k = k + 1 
  } 
  m/d(x)
}

w = function(b, v, q = T){
  n = d(b)
  o = 0
  for(i in seq_len(n)){ 
    o = o + s(b[i]-v) 
  } 
  (1/(n-1*(q == T))*o)^(0.5)
}

n = function(p, i, ...){ m = u(i, ...); r = w(i, m); (p-m)/r}

z = function(p, i, h = T, ...){
  g = seq(-3,3,length=1000);gg = dnorm(g)
  plot(g, gg, type="l", lwd=1, ...)
  k = n(p = p,i = i, ...)
  if(h) { tt = c(g[g<=k],k,-3); ll = c(gg[g<=k],0,0) }else{ tt = c(k,g[g>=k],3); ll = c(0,gg[g>=k],0) }
  polygon(tt, ll, col="red")
}
```
4. Make sure the output is the **same** between your version and the cleaned up version.


    - You may wish to individually test each function using `all.equal(bad_func(),your_func())`.
  
## (5 Points) Exercise 3 (Recursion is Recursion is Recursion is ...)

> "To iterate is human, to recurse divine."
>
> --- L. Peter Deutsch

Within this exercise, you will explore the concept of recursion and converting to and fro *tail* recursive statements. 

A common operation within Statistics is to obtain an **average** over data. The average in this case is given as: \[{ {\bar x}_n} = \frac{1}{n}\sum\limits_{i = 1}^n { {x_i} } \] where the subscript $n$ on ${ {\bar x} }$ represents the sample size.

- Implement two functions that obtains the average of a vector `x` where one function uses a loop and the other is vectorized.
    - The function must return `NaN` if the length of `x` is `0`. 
    - If the vector `x` contains `NA` values, then remove them using `na.omit()`.
    - **Note:** You are *not* permitted to call the base R implementation of `mean()`. However, you should check if the values are equal using `all.equal(x,y)`.

```{r}
x_bar_loop = function(x) {
    if (length(na.omit(x)) == 0) {
    mean1 = NaN  
  } else {
    mean1 = 0  
    for (i in 1: length(na.omit(x))) {
      mean1 = mean1 + (na.omit(x)[i])/(length(na.omit(x)))
    }
  }
  return(mean1)
}
```

```{r}
x_bar_vector = function(x) {
  if (length(na.omit(x)) == 0) {
    mean2 = NaN
  } else {
    mean2 = (sum(na.omit(x))/length(na.omit(x)))
  }
  return(mean2)
}
```

- We can also implement a recursive cummulative average if we modify the definition for ${ {\bar x}_n}$ to depend on itself by ${ {\bar x}_{n-1} }$. 
    - Try to show that $${ {\bar x}_n} = \frac{1}{n}\left[ {\left( {n - 1} \right){ {\bar x}_{n - 1} } + {x_n} } \right]$$.

- Please implement the following recursive formula:
\[ { {\bar x}_n} = \frac{1}{n}\left[ {\left( {n - 1} \right){ {\bar x}_{n - 1} } + {x_n} } \right]\]
    - Do not worry about returning `NaN` or detecting `NA` values. Focus only on writing the recursion case.

```{r}
x_bar_recurse = function(x) {
  n = length(x)
  if (n == 1) {
    return(x[1])
  } else {
    return((((n-1)*x_bar_recurse(x[-n]) + x[n])/n))
  }
}
```

## (5 Points) Exercise 4 (Bugs, Bugs Everywhere!)

One of the many inveitiable side effects of programming is bugs. As Alexander so eloquently put it:

> "99 Little Bugs in the Code, take one down, patch it around, 117 little bugs in the code"
> 
> - Alexander via [Twitter](https://twitter.com/irqed/status/358212928404586498)

The next section contains many "errors" that appear in R code. These errors are actual cases that I have encountered while working with various students or on different projects. Your job is to solve them and explain *how* you fixed the problem. 

   1. Specify what the error is that you encounter when running the code.
   1. Briefly explain the implemented solution to fix the function.
   1. **One Caveat:** You are *not* allowed to modify the data that goes into the function. (Modifying the internals of functions is okay!)
   
- Why does my function not output anything!??!?
```{r infunc, eval = FALSE}
f = function(x){
  n = a+x
}

f(1)
```
**The function contains object (a) that was not defined before. To fix the bug, we could define `a` first. also we should print the output if we want show it.**
```{r infunc, eval = FALSE}
f = function(x){
  a = x+1 #define `a` here
  n = a+x
}

(f(1)) #print
```

- Huh, I could of sworn the function was working a minute ago. I made sure that the calculation were vectorized. Why doesn't it work!?
```{r, eval = FALSE}
p = function(x){
  log(p/(1-p))
}

vals = seq(0, 1, by = 0.1)

p(vals)
```
**p is a function, 1-p couldn't be calculated. we could `log(p/(1-p)` to `log(p(x)/(1-p(x))`.**

- The mean seems to be broken...
```{r, eval = FALSE}
d = c(1, 3, NA, 5, 6, 9)

mean(d) == 1/length(d)*sum(d)

```
**we should use there is a missting value in `d`, we should replace `d` into `na.omit(d)` in the formula. We might also check whether we should use `=` or `==` depending on the object of the formula.**

- If this, then that right? 
```{r, eval = FALSE}
set.seed(145)

d = sample(1:10, 5)
b = sample(1:10, 5)

if(d > b){
  d+5
}
else {
  b*2
}
```
**Use `ifelse(d > b, d+5, d*2)` instead.**

- What is this blasphemy?! Why am I not getting any successful runs?? 
```{r, eval = FALSE}
set.seed(1234)

f = function() {
    if (runif(1) > .8) stop("oops")
    TRUE
}

g = function() {
  tryCatch(f(), 
           error = function(err) { warning(conditionMessage(err)) }
           )
}

a = replicate(10, g())
if(is.logical(a) && sum(a) > 0){
  cat("We had a few successes!\n")
}
```


## (5 Points) Exercise 5 (Memoize Me!)

The objective here is to construct different functions that are able to perform underneath memoization.

Recall from [Lecture 5](http://stat385.thecoatlessprofessor.com/assets/lectures/lec5/lec5_functions_recursion_memoization_benchmarking.pdf#page=37) the `factorial`.

- Create one *internal* and one *external* **memoized** factorial function.
  - Do **not** name it `factorial()`.
  
```{r, cache = T}
install.packages("memoise")
```


```{r}
fac_r <- function(x){ 
if(x <= 1){
return(1)
} else {
return(x*fac_r(x-1))
}
}
```

```{r}
library("memoise")
ex_mem_fac = memoise(fac_r)
in_mem_fac = memoise(function(x) {
  if(x <= 1){
return(1)
} else {
return(x*in_mem_fac(x-1))
}
})

fac_basic = function(x) {
  return(factorial(x))
}
```

- Check for equality between the factorial functions using the base R implementation of `factorial()`.

```{r}
all.equal(ex_mem_fac, in_mem_fac, fac_basic)
```

- Perform a benchmark on all *four* functions. 
  - Compare the function against the base R and the two memoized versions.
```{r, cache = T}
set.seed(1281)
x = sample(50:100,1)
```

```{r, cache = T}
forget(ex_mem_fac)
forget(in_mem_fac)
forget(fac_r)
forget(fac_basic)
fac_r <- function(x){ 
if(x <= 1){
return(1)
} else {
return(x*fac_r(x-1))
}
}
system.time({fac_r(x)})
```

```{r, cache = T}
forget({fac_r})
ex_mem_fac = memoise(fac_r)
system.time({ex_mem_fac(x)})
system.time({ex_mem_fac(x)})
(ex_mem_fac(x))
```

```{r, cache = T}
forget(ex_mem_fac)
forget(fac_r)
in_mem_fac = memoise(function(x) {
  if(x <= 1){
return(1)
} else {
return(x*in_mem_fac(x-1))
}
})
system.time({in_mem_fac(x)})
system.time({in_mem_fac(x)})
(in_mem_fac(x))
```

```{r, cache = T}
forget(ex_mem_fac)
forget(in_mem_fac)
forget(fac_r)
fac_basic = function(x) {
  return(factorial(x))
}
system.time({fac_basic(x)})
system.time({fac_basic(x)})
(fac_basic(x))
```

- Use the function with the best *timing* to implement combinations: $$C\left( {n,k} \right) = \frac{ {n!} }{ {k!\left( {n - k} \right)!} }$$
  - Compare the combination function against `choose()` in base R. 

**the external cache is more effective, so we use external cache**
```{r}
set.seed(1228)
n = sample(20:30, 1)
k = sample(10:20, 1)
combination = function(n, k) {
  return((ex_mem_fac(n))/((ex_mem_fac(k))*(ex_mem_fac(n-k))))
}
```

```{r, cache = T}
all.equal(combination(n, k), choose(n,k))
```

